{\fontfamily{ptm}\selectfont Nesterov Type Momentum Methods}

### 2024-05-28 Notebook LaTEX/Momentum Accelerations/main.tex
>w< mbp

\section{The Proximal Point Method in the Convex Case}
    In this section we quickly go over the analysis of Proximal point method (PPM) in the convex case and see how the theories can be generalized into the cases where PPM is approximated. 
    \par
    With $f: \RR^n \mapsto \overline \RR$ lsc proper and convex, 
    given any $x_0$ the PPM generates sequence $(x_n)_{n\in \NN}$ by $x_{k + 1} = \hprox_{\eta_{k + 1}}f(x_k)$ for all $k \in \NN$ where the sequence $(\eta_{k})_{k \in \NN}$ is a nonegative sequence of real numbers. 
    \subsection{The Lyapunov function of PPM}
        \begin{theorem}
            Define the quantity for all $u \in \RR^n$
            $$
                \Phi_t = 
                \left(
                    \sum_{i = 1}^{t} \eta_i
                \right)(f(x_t) - f(u)) + \frac{1}{2} \Vert u - x_t\Vert^2 \quad \forall t \ge 1, 
            $$
            then it is a Lyapunov function for the PPM algorithm. 
            Meaning for all $(x_k)_{k \in \NN}$ generated by PPM, it satisfies that $\Phi_{t + 1} - \Phi_t \le 0$. 
            Additionally, by the definition we have 
            \begin{align*}
                \Phi_{t + 1} - \Phi_{t} = 
                \eta_{t + 1}(f(x_{t + 1}) - f(u)) 
                + 
                \frac{1}{2}\Vert u - x_{t + 1}\Vert^2 
                -
                \frac{1}{2}\Vert u - x_t\Vert^2 
                \le 
                - \frac{1}{2}\Vert x_{t + 1} - x_t\Vert^2, 
            \end{align*}
            as a consequence of choosing $u = x_{t + 1}$
            \begin{align*}
                f(x_{t + 1}) - f(x_t) \le \frac{1}{\eta_{t + 1}}\Vert x_{t + 1} - x_t\Vert^2. 
            \end{align*}
        \end{theorem}
        \begin{proof}
            With $p, \alpha, y$ in theorem \ref*{thm:prox_descent_ineq} as $x_{t + 1}, \eta_{t + 1}, u$ we have for all $u$
            \begin{align*}
                \eta_{t + 1} (f(x_{t + 1}) - f(u)) + 
                \frac{1}{2}(
                    \Vert x_t - x_{t + 1} \Vert^2 
                    - 
                    \Vert x_t - u\Vert^2
                ) &\le 
                - \frac{1}{2}\Vert u - x_{t + 1}\Vert^2
                \\
                \eta_{t + 1}(f(x_{t + 1}) - f(u)) 
                + 
                \frac{1}{2}\Vert u - x_{t + 1}\Vert^2 
                -
                \frac{1}{2}\Vert u - x_t\Vert^2 
                & \le 
                - \frac{1}{2}\Vert x_{t + 1} - x_t\Vert^2, 
            \end{align*}
            next it's not hard to check that $\Phi_{t + 1} - \Phi_t$ equals to the above quantity. 
        \end{proof}
        \begin{remark}
            This theorem act as a descent inequality and it can be generalized under a diverse context, such as all algorithms that are approximation to the PPM method. 
        \end{remark}

\section{Application of the Proximal Point Method}

>>>>

\section{The Proximal Point Method in the Convex Case}
    In this section we quickly go over the analysis of Proximal point method (PPM) in the convex case and see how the theories can be generalized into the cases where PPM is approximated. 
    \par
    With $f: \RR^n \mapsto \overline \RR$ lsc proper and convex, 
    given any $x_0$ the PPM generates sequence $(x_n)_{n\in \NN}$ by $x_{k + 1} = \hprox_{\eta_{k + 1}}f(x_k)$ for all $k \in \NN$ where the sequence $(\eta_{k})_{k \in \NN}$ is a nonegative sequence of real numbers. 
    \subsection{The Lyapunov function of PPM}
        \begin{theorem}
            Define the quantity for all $u \in \RR^n$
            \begin{align*}
                \Phi_t &= 
                \left(
                    \sum_{i = 1}^{t} \eta_i
                \right)(f(x_t) - f(u)) + \frac{1}{2} \Vert u - x_t\Vert^2 \; \forall t \ge 1, 
                \\
                \Phi_0 &= (1/2)\Vert x_0 - u\Vert^2, 
            \end{align*}
            then it is a Lyapunov function for the PPM algorithm. 
            Meaning for all $(x_k)_{k \in \NN}$ generated by PPM, it satisfies that $\Phi_{t + 1} - \Phi_t \le 0$. 
            Additionally, by the definition we have 
            \begin{align*}
                \Phi_{t + 1} - \Phi_{t} = 
                \eta_{t + 1}(f(x_{t + 1}) - f(u)) 
                + 
                \frac{1}{2}\Vert u - x_{t + 1}\Vert^2 
                -
                \frac{1}{2}\Vert u - x_t\Vert^2 
                \le 
                - \frac{1}{2}\Vert x_{t + 1} - x_t\Vert^2, 
            \end{align*}
            as a consequence of choosing $u = x_{t + 1}$
            \begin{align*}
                f(x_{t + 1}) - f(x_t) \le \frac{1}{\eta_{t + 1}}\Vert x_{t + 1} - x_t\Vert^2. 
            \end{align*}
        \end{theorem}
        \begin{proof}
            With $p, \alpha, y$ in theorem \ref*{thm:prox_descent_ineq} as $x_{t + 1}, \eta_{t + 1}, u$ we have for all $u$
            \begin{align*}
                \eta_{t + 1} (f(x_{t + 1}) - f(u)) + 
                \frac{1}{2}(
                    \Vert x_t - x_{t + 1} \Vert^2 
                    - 
                    \Vert x_t - u\Vert^2
                ) &\le 
                - \frac{1}{2}\Vert u - x_{t + 1}\Vert^2
                \\
                \eta_{t + 1}(f(x_{t + 1}) - f(u)) 
                + 
                \frac{1}{2}\Vert u - x_{t + 1}\Vert^2 
                -
                \frac{1}{2}\Vert u - x_t\Vert^2 
                & \le 
                - \frac{1}{2}\Vert x_{t + 1} - x_t\Vert^2, 
            \end{align*}
            next it's not hard to check that $\Phi_{t + 1} - \Phi_t$ equals to the above quantity. 
        \end{proof}
        \begin{remark}
            This theorem act as a descent inequality and it can be generalized under a diverse context, such as all algorithms that are approximation to the PPM method. 
        \end{remark}

        \begin{theorem}[Convergence Rate of PPM]
            The convergence rate of PPM applied to $f$, closed, convex proper, we have convergence rate of the function value: 
            $$
            f(x_T) - f(x_*) \le O\left(\left(\sum_{i=1}^{T}\eta_t\right)^{-1}\right). 
            $$
            Where $x_*$ is the minimizer of $f$. 
        \end{theorem}
        \begin{proof}
            With $\Delta_t = f(x_t) - f(x_*), \Upsilon_t = \sum_{i = 1}^{t}\eta_i$ so $\Phi_t = \Upsilon_t\Delta_t + \frac{1}{2}\Vert x_t - x_*\Vert^2$ by consider $u = x_*$, invoking previous theorem and do
            \begin{align*}
                \Upsilon_T\Delta_T \le \Phi_T 
                &\le 
                \Phi_0 = \frac{1}{2}\Vert x_0 - x_*\Vert^2 
                \\
                \implies \Delta_T 
                &\le 
                \frac{1}{2\Upsilon_T} \Vert x_0 - x_*\Vert^2. 
            \end{align*}
            Hence, the convergence rate of PPM for all convex function is the above. 
        \end{proof}
        \begin{remark}
            The analysis of the above is taken from (REFERENCE NEEDED). 
        \end{remark}
\section{Application of the Proximal Point Method}
    The PPM method and the Lyaounov function derived above serves as tamplate for other algorithms. 
    In optimizations, people uses lower and upper approximation of the objective function to approximate the PPM. 
    Such an approach involves a diverse range of algorithms, including second order algorithms such as Newton's method. 
    To demonstrate, assume that $f$ is a lsc convex function such that it can be approximated by an lower bounding function $l_f(x|\bar x)$ at $\bar x$ such that it satisfies for all $x$: 
    \begin{align}
        l_f(x| \bar x) 
        \le f(x) \le l_f(x|\bar x) + \frac{L}{2}\Vert x - \bar x\Vert^2. 
    \end{align}
    The above characterization is generic enough to include the case where $l_f(x|\bar x)$ under approximates function that is non-smooth. 
    To make use of the theorems discussed previously, we assume that $l_f(x|\bar x)$ is convex for all $x$, at all $\bar x$. 
    \par 
    The approximated proximal point method is applying PPM to the function $l_f(x|x_t)$ for each iteration, i.e: $x_{t +1} = \hprox_{\eta_{t + 1}}[u\mapsto l_f(u | x_t)](x_t)$.
